---
title:"Optimizing Portfolio"
author: "ROHIT THAKUR"
PGID: "12120040"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: inline
---



```{r}
rm(list=ls())
pacman::p_load(tidyverse,tidyquant,PortfolioAnalytics,quantmod,PerformanceAnalytics,zoo,
               plotly,ggthemes,timetk)
library(PortfolioAnalytics)
library(crayon)
library(quantmod)

```

```{r}

#SBI BANK
SBIBANK <- tseries::get.hist.quote(instrument="SBIN.NS", start="2007-07-01",
                      end="2018-07-31", quote="AdjClose",
                      provider="yahoo", origin="1970-01-01",
                      compression="m", retclass="zoo")
SBIBANK <- na.omit(SBIBANK)

#TATA MOTORS
TATAMOTORS <- tseries::get.hist.quote(instrument="TATAMOTORS.NS",start="2007-07-01",
                      end="2018-07-31", quote="AdjClose",
                      provider="yahoo", origin="1970-01-01",
                      compression="m", retclass="zoo")
TATAMOTORS <- na.omit(TATAMOTORS)
#HCLTECH
HCLTECH <- tseries::get.hist.quote(instrument="HCLTECH.NS", start="2007-07-01",
                      end="2018-07-31", quote="AdjClose",
                      provider="yahoo", origin="1970-01-01",
                      compression="m", retclass="zoo")
HCLTECH <- na.omit(HCLTECH)
Britannia <- tseries::get.hist.quote(instrument="BRITANNIA.NS", start="2007-07-01",
                      end="2018-07-31", quote="AdjClose",
                      provider="yahoo", origin="1970-01-01",
                      compression="m", retclass="zoo")
Britannia <- na.omit(Britannia)
DLF <- tseries::get.hist.quote(instrument="DLF.NS", start="2007-07-01",
                      end="2018-07-31", quote="AdjClose",
                      provider="yahoo", origin="1970-01-01",
                      compression="m", retclass="zoo")
DLF <- na.omit(DLF)
HeroMotors <- tseries::get.hist.quote(instrument="HEROMOTOCO.NS", start="2007-07-01",
                      end="2018-07-31", quote="AdjClose",
                      provider="yahoo", origin="1970-01-01",
                      compression="m", retclass="zoo")
HeroMotors <- na.omit(HeroMotors)
IOC <- tseries::get.hist.quote(instrument="IOC.NS", start="2007-07-01",
                      end="2018-07-31", quote="AdjClose",
                      provider="yahoo", origin="1970-01-01",
                      compression="m", retclass="zoo")
IOC <- na.omit(IOC)
AXIS <- tseries::get.hist.quote(instrument="AXISBANK.NS", start="2007-07-01",
                      end="2018-07-31", quote="AdjClose",
                      provider="yahoo", origin="1970-01-01",
                      compression="m", retclass="zoo")
AXIS <- na.omit(AXIS)
ZeeTV <- tseries::get.hist.quote(instrument="ZEEL.NS", start="2007-07-01",
                      end="2018-07-31", quote="AdjClose",
                      provider="yahoo", origin="1970-01-01",
                      compression="m", retclass="zoo")
ZeeTV <- na.omit(ZeeTV)
TCS <- tseries::get.hist.quote(instrument="TCS.NS", start="2007-07-01",
                      end="2018-07-31", quote="AdjClose",
                      provider="yahoo", origin="1970-01-01",
                      compression="m", retclass="zoo")
TCS <- na.omit(TCS)

#MERGING ALL THE STOCK PRICES      
Portfolio.Prices <- merge(SBIBANK, TATAMOTORS,HCLTECH,Britannia,DLF,
                      HeroMotors, IOC,AXIS,ZeeTV,TCS)   
  
Portfolio.Returns <- NULL

#CALCULATING THE RETURNS FOR THE PORTFOLIO
Portfolio.Returns = Return.calculate(Portfolio.Prices, method="simple")
Portfolio.Returns = na.omit(Portfolio.Returns)


Portfolio.Names <- c("SBIBANK", "TATAMOTORSitals", "HCLTECH", "Britannia", "DLF",
                 "HeroMotors", "IOC", "AXIS", "ZeeTV", "TCS")

#Define column names 
names(Portfolio.Returns) <- Portfolio.Names

# Set Equal weight benchmark
n <- ncol(Portfolio.Returns)
equal_weights <- rep(1 / n, n)


benchmark_returns <- Return.portfolio(R = Portfolio.Returns,
                                      weights = equal_weights, 
                                      rebalance_on = "years")

colnames(benchmark_returns) <- "Benchmark"

# Benchmark - Equal weightage Portfolio
table.AnnualizedReturns(benchmark_returns)
# Given equal weights of 0.1 for 10 assets
```

```{r}
cat(bold(red('Objectives for Optimization\n')))

# 1. minimize variance
# 2. maximize return
# 3. maximize quadratic utility
# 4. minimize ETL

# Create portfolio specification object
pspec <- portfolio.spec(assets=Portfolio.Names)

cat(bold(red('DEFINE CONSTRAINTS AND OBJECTIVES\n')))

# Full investment constraint
FI_Constr <- weight_sum_constraint(min_sum=1, max_sum=1)

# Long only constraint
LO_Constr <- box_constraint(assets=pspec$assets, min=0, max=1)

# Position limit constraint
PosLimit_Constr <- position_limit_constraint(assets=pspec$assets, max_pos=3)

# Group constraint
Group_Constr <- group_constraint(assets=pspec$assets, groups=list(c(1,5,8,10), c(2,3,4,6), c(7,9)), 
                                 group_min=0, group_max=0.5)

# Position limit constraint
PosLimit_Constr <- position_limit_constraint(assets=pspec$assets, max_pos=3)

# Target mean return constraint
TargetReturn_constr <- return_constraint(return_target=0.015)

# Diversification constraint
Div_Constr <- diversification_constraint(div_target=0.7)

# Turnover constraint
TurnOver_Constr <- turnover_constraint(turnover_target=0.2)

# Transaction cost contraint
TransCost_Constr <- transaction_cost_constraint(assets=pspec$assets, ptc=0.01)

##### Objectives #####

# Return objective
maxret_obj <- return_objective(name="mean")

# Risk objective
minrisk_obj <- portfolio_risk_objective(name="var")

# ETL objective
etl_obj <- portfolio_risk_objective(name="ETL")

#***************1. Min Variance Optimization**********************

# FOR Global minimum variance portfolio, Only specify the full investment constraint and min risk objective
library(ROI)
library(quadprog)
library(ROI.plugin.quadprog)
library(ROI.plugin.glpk)
library(ROI.plugin.symphony)
library(PerformanceAnalytics)
# Consider Global minimum portfolio as Base portfolio
opt_base <- optimize.portfolio(R=Portfolio.Returns, portfolio=pspec,
                                 constraints=list(FI_Constr),
                                 objectives=list(minrisk_obj),
                                 optimize_method="ROI")
opt_base

```
```{r}
# Calculate base portfolio returns
base_returns <- Return.portfolio(Portfolio.Returns, extractWeights(opt_base))

colnames(base_returns) <- "Base"
####Global min portfolio back testing

#Run the optimization with periodic balancing
bt_opt_base <- optimize.portfolio.rebalancing(R = Portfolio.Returns, portfolio = pspec,
                                           constraints = list(FI_Constr), 
                                           objectives = list(minrisk_obj), 
                                           rebalance_on = "months", optimize_method = "ROI",
                                           training_period = 60, rolling_window = 60)

backtesting_base_returns <- Return.portfolio(Portfolio.Returns, extractWeights(bt_opt_base))

colnames(backtesting_base_returns) <- "BackTesting_Base"

# Annualized performance
table.AnnualizedReturns(cbind(benchmark_returns, base_returns, backtesting_base_returns))
#Back Testing yielded better results in the past. so our base portfolio is good.
```

```{r}

# Chart the optimal weights for the back tested base portfolio 
chart.Weights(bt_opt_base, colorset=tim12equal)

#Tatamotors and DLF has got negative weightages when forming global min portfolio

```


```{r}
#######CONSTRAINED MIN VARIANCE WITH ROI############

#Min variance portfolio with target return constraint
tr_opt_minvar <- optimize.portfolio(R=Portfolio.Returns, portfolio=pspec,
                                 constraints = list(FI_Constr, LO_Constr,
                                                    TargetReturn_constr),
                                 objectives = list(minrisk_obj),
                                 optimize_method = "ROI", trace=TRUE)

tr_opt_minvar

```

```{r}
#Calculate returns under specified constraints and objectives
tr_minvar_returns <- Return.portfolio(Portfolio.Returns, extractWeights(tr_opt_minvar))
colnames(tr_minvar_returns) <- "Min Variance with TR"

# Backtest
bt_opt_minvar <- optimize.portfolio.rebalancing(R = Portfolio.Returns, portfolio = pspec,
                                             constraints = list(FI_Constr,LO_Constr,
                                                                TargetReturn_constr),
                                             objectives = list(minrisk_obj),
                                             rebalance_on = "months", optimize_method = "ROI",
                                             training_period = 60, rolling_window = 60)

bt_minvar_returns <- Return.portfolio(Portfolio.Returns, extractWeights(bt_opt_minvar))

colnames(bt_minvar_returns) <- "With Back Testing"

table.AnnualizedReturns(cbind(tr_minvar_returns, bt_minvar_returns))
#Backtesting the min variance portfolio when target return is set to 0.15, 
#shows that not only the realized return is slightly higher than the expected return 
#but also actual risk might be greater than the expected risk.
#However we can recommend min variance portfolio of 15% TR as it has higher relaized sharpe ratio.
```

```{r}
chart.Weights(bt_opt_minvar, colorset=tim12equal)

```

```{r}
#Box constraint 1 - a common cap for all the assets
Box_Constr1 <- box_constraint(assets=pspec$assets, min=0.01, max=0.4)


box1_opt_minvar <- optimize.portfolio(R=Portfolio.Returns, portfolio=pspec,
                                      constraints = list(FI_Constr, LO_Constr,
                                                         TargetReturn_constr, Box_Constr1),
                                      objectives = list(minrisk_obj),
                                      optimize_method = "ROI", trace=TRUE)
box1_opt_minvar

cat("Box Constraint1 - Min:", Box_Constr1$min[[1]], " Max: ", Box_Constr1$max[[1]], "\n")
```

```{r}
#Calculate returns under specified constraints and objectives
box1_minvar_returns <- Return.portfolio(Portfolio.Returns, extractWeights(box1_opt_minvar))
colnames(box1_minvar_returns) <- "MinVariance_TR_Box1"

# Backtest for min variance with TR and BOX2 constraints
bt_box1_opt_minvar <- optimize.portfolio.rebalancing(R = Portfolio.Returns, portfolio = pspec,
                                             constraints = list(FI_Constr,LO_Constr,
                                                                TargetReturn_constr, Box_Constr1),
                                             objectives = list(minrisk_obj),
                                             rebalance_on = "months", optimize_method = "ROI",
                                             training_period = 60, rolling_window = 60)

bt_box1_minvar_returns <- Return.portfolio(Portfolio.Returns, extractWeights(bt_box1_opt_minvar))
colnames(bt_box1_minvar_returns) <- "BackTesting_TR_Box1"

table.AnnualizedReturns(cbind(box1_minvar_returns, bt_box1_minvar_returns))
##Putting cap on the asset weightage yielded higher return than the expected return 
#but with a slightly increased risk and lower sharpe ratio.
#We can still experiment with different min and max values to see if we can get better optimal weights
#and there by higher returns to risk for our portfolio.
```

```{r}
chart.Weights(bt_box1_opt_minvar, colorset=tim12equal)

```

```{r}
#Box constraint 2 - separate caps for individual assets
Box_Constr2 <- box_constraint(assets = pspec$assets,
                            min=c(0, 0.02, 0.02, 0.04, 0.01, 0.03, 0.04, 0.01, 0.005, 0),
                            max=c(0, 0.7, 0.8, 0.6, 0.8, 0.75, 0.8, 0.7, 0.8, 0))
box2_opt_minvar <- optimize.portfolio(R=Portfolio.Returns, portfolio=pspec,
                                      constraints = list(FI_Constr, LO_Constr,
                                                         TargetReturn_constr, Box_Constr2),
                                      objectives = list(minrisk_obj),
                                      optimize_method = "ROI", trace=TRUE)
box2_opt_minvar
```
```{r}
#Calculate returns under specified constraints and objectives
box2_minvar_returns <- Return.portfolio(Portfolio.Returns, extractWeights(box2_opt_minvar))
colnames(box2_minvar_returns) <- "MinVariance_TR_Box2"
table.AnnualizedReturns(cbind(box1_minvar_returns, box2_minvar_returns))
#Box 2 constraint on min variance portfolio has better return to risk ratio compared to 
#Box 1 constraint. Which means rather than generalizing same cap values for all the assets,
# we should have individual cap values on each asset according to thier historical return data.
```


```{r}
#add box and group constraint but no target return
bg_opt_minvar <- optimize.portfolio(R = Portfolio.Returns, portfolio = pspec,
                                                constraints = list(FI_Constr, LO_Constr, 
                                                             Box_Constr2, Group_Constr),
                                                optimize_method = "ROI", trace=TRUE)
bg_opt_minvar

```

```{r}
bg_minvar_returns <- Return.portfolio(Portfolio.Returns, extractWeights(bg_opt_minvar))
colnames(bg_minvar_returns) <- "MinVarBox&Group"

bt_bg_opt_minvar <- optimize.portfolio.rebalancing(R = Portfolio.Returns, portfolio = pspec,
                                          constraints = list(FI_Constr, LO_Constr, 
                                                             Box_Constr2, Group_Constr),
                                          rebalance_on = "months", optimize_method = "ROI",
                                          training_period = 60, rolling_window = 60)

# Calculate portfolio returns
bt_bg_minvar_returns <- Return.portfolio(Portfolio.Returns, extractWeights(bt_bg_opt_minvar))
colnames(bt_bg_minvar_returns) <- "BackTestingBox&Group"
# Annualized performance 
table.AnnualizedReturns(cbind(bg_minvar_returns, bt_bg_minvar_returns))
#Portfolio with box and group constraints indiactes high risk than returns
#Box2 and group constraints combined together did not yield better results. 
#so portfolio with these constraitns is not recommended for this case.
```

```{r}
compareReturns <- cbind(base_returns, tr_minvar_returns, 
                        box1_minvar_returns, box2_minvar_returns, bg_minvar_returns)
colnames(compareReturns) <- c("Base", "MinVar_TR", "MinVar_Box1", "Minvar_Box2", "Minvar_Box2_Group")

# compare portfolios with different constraints
table.AnnualizedReturns(compareReturns)
#Our base portfolio has the best return to risk ratio compared to min variance portfolio with constraints.
```

```{r}
#Create an initial portfolio object with leverage and box constraints
init_portfolio <- portfolio.spec(assets=Portfolio.Names)
init_portfolio <- add.constraint(portfolio=init_portfolio, type="leverage", min_sum=0.99, max_sum=1.01)
init_portfolio <- add.constraint(portfolio=init_portfolio, type="box", min=0.05, max=0.65)

#Add constraints to maximize mean returns
maxreturns <- add.objective(portfolio=init_portfolio, type="return", name="mean")
#Run the optimization.
MaxReturns_OptPortfolio <- optimize.portfolio(R=Portfolio.Returns, portfolio=maxreturns,
                                              optimize_method="ROI", trace=TRUE)
MaxReturns_OptPortfolio

```

```{r}
chart.RiskReward(MaxReturns_OptPortfolio,return.col="mean", risk.col="StdDev",
 chart.assets=TRUE, main="Risk Reward Chart for Maximum Return")
```
```{r}
#Back Testing
#Initil training period of 60 months is chosen
bt_maxret <- optimize.portfolio.rebalancing(Portfolio.Returns, maxreturns,
                                            optimize_method="ROI", rebalance_on="months",
                                            training_period=60)
bt_maxret


```
```{r}
chart.Weights(bt_maxret, colorset=tim12equal)
```
```{r}
max_returns <- Return.portfolio(Portfolio.Returns, extractWeights(MaxReturns_OptPortfolio))
colnames(max_returns) <- "Max Return Portfolio"

bt_max_returns <- Return.portfolio(Portfolio.Returns, extractWeights(bt_maxret))
colnames(bt_max_returns) <- "BackTesting"

table.AnnualizedReturns(cbind(max_returns, bt_max_returns))
#We can recommend Max Return Portfolio as the back testing shows that the realized returns are higher
#with significantly lower risk compared to expected returns and risk values.
```


```{r}

cat(bold(red('Maximizing Quadratic Utility\n')))

#Set constraints - full investment, long only
#set objectives - return, risk and a risk aversion value of 0.35

ret_obj <- return_objective(name="mean")
# risk_aversion parameter is for controlling how much the variance need to be penalized
var_obj <- portfolio_risk_objective(name="var", risk_aversion=0.35)
#Run the optimization.
opt_qu <- optimize.portfolio(R=Portfolio.Returns, portfolio=pspec, 
                             constraints = list(FI_Constr, LO_Constr),
                             objectives = list(ret_obj, var_obj),
                             optimize_method='ROI', trace=TRUE)
print(opt_qu)

#Except for Britannia and HCL Tech all the other assets are given a weightage of 0 to maximize the utility.
#Note: Risk aversion is set to 0.35 because we got all NA values when risk aversion is set to 0.25
```


````{r}
#BackTesting
bt_qu <- optimize.portfolio.rebalancing(Portfolio.Returns, portfolio=pspec,
                                        constraints=list(FI_Constr, LO_Constr), 
                                        objectives=list(ret_obj, var_obj),
                                        optimize_method='ROI', rebalance_on='months',
                                        training_period=60, rolling_window = 60)
bt_qu <- na.omit(bt_qu)
qu_returns <- Return.portfolio(Portfolio.Returns, na.omit(extractWeights(opt_qu)))
colnames(qu_returns) <- "QU"

bt_qu_returns <- Return.portfolio(Portfolio.Returns, na.omit(extractWeights(bt_qu)))
colnames(bt_qu_returns) <- "BackTested_QU"

table.AnnualizedReturns(cbind(qu_returns, bt_qu_returns))
#We can recommend Max QU portfolio because Back testing this portfolio has better return to risk ratio, less deviation/risk and realized returns amost match with the expected returns.
```

```{r}
#Minimizing Expected Tail Loss

etl_Box_constr <- box_constraint(assets=pspec$assets, min=0.05, max=0.65)
leverage_constr <- leverage_exposure_constraint(assets=pspec$assets, min_sum=0.99, max_sum=1.01)

opt_etl <- optimize.portfolio(Portfolio.Returns, portfolio=pspec, 
                              constraints = list(etl_Box_constr, leverage_constr),
                              objectives = list(etl_obj),
                              optimize_method = "ROI", trace=TRUE)
opt_etl

```
```{r}
chart.RiskReward(opt_etl,return.col="mean", risk.col="StdDev",
 chart.assets=TRUE, main="Risk Reward Chart for Mimimizing Expected Tail Loss")
```
```{r}
etl_returns <- Return.portfolio(Portfolio.Returns, extractWeights(opt_etl))
colnames(etl_returns) <- "ETL"

bt_opt_etl <- optimize.portfolio.rebalancing(Portfolio.Returns, portfolio = pspec,
                                             constraints = list(etl_Box_constr, leverage_constr),
                                             objectives = list(etl_obj),
                                             optimize_method = "ROI", rebalance_on = "months",
                                             training_period = 60, rolling_window = 60)

bt_etl_returns <- Return.portfolio(Portfolio.Returns, extractWeights(bt_opt_etl))
colnames(bt_etl_returns) <- "BackTested_ETL"

table.AnnualizedReturns(cbind(etl_returns, bt_etl_returns))

#We can recommed Min ETL portfolio as the back tested results indicate better return to risk ratio. 
#Back tested reults are less deviating compared to expected results eventhough the actual return is 2% less than the expected return.

```
finalComparision <- table.AnnualizedReturns(cbind(base_returns, tr_minvar_returns, box1_minvar_returns, 
                         max_returns, qu_returns, etl_returns))

colnames(finalComparision) <- c("Base", "MinRisk", "Box1", "MaxReturn", "QU", "ETL")

backtestedFinal <- table.AnnualizedReturns(cbind(backtesting_base_returns, bt_minvar_returns, 
                                                bt_box1_minvar_returns, bt_max_returns, 
                                                bt_qu_returns, bt_etl_returns ))

colnames(backtestedFinal) <- c("Base", "MinRisk", "Box1", "MaxReturn", "QU", "ETL")

#final portfolio results
finalComparision

```

```{r}
#Back tested results
backtestedFinal
#Back testing consistently showed that all the portfolios have better returns in the past except for the 
#portfolio with box1 constraint.
```

